Identifying covariate traits of trance and possession phenomena in the Ethnographic

### 2022-10-20 Readme.Rmd
some more work

There are eight types of trance and possession phenomena recorded in the Atlas and a wide set of possible covariates, all with a number of levels.

```{r }
vars = ld %>% 
  filter(var_title != 'Trance states') %>% 
  distinct(var_id,var_title,var_name) %>%
  arrange(var_id) %>% 
  group_by(var_id, var_title) %>% 
  nest() %>% 
  mutate(levels = map(data, ~ paste(.$var_name, collapse = ', '))) %>% 
  select(var_id, var_title, levels) %>% 
  unnest(cols = c(levels)) %>% 
  rename(ID = var_id, Name = var_title)

kable(vars, 'simple')
```

@@

Following @bourguignon1968, we reduce the complexity of the outcome variable to two levels: whether a society has possession trance or not. The predictors are the possible covariates identified above. Some predictors, like `EA032` (Jurisdictional hierarchy of local community) or `EA113` can be construed to follow a scale. These are input as ordered factors. Others, like `EA053` (Sex differences in animal husbandry) or `EA043` (Major type of descent) can be carved up into categories: These are one-hot encoded. For details, see the helper file.

@@

```{r }
varimp1 = read_tsv('models/varimp11.tsv')
vars2 = varimp1 %>% 
  slice(1:10) %>% 
  mutate(ID = str_extract(variable, '^EA...')) %>% 
  left_join(vars, by = "ID") %>% 
  select(ID, Name, relative_importance)

@@

>>>>

There are eight types of trance and possession phenomena recorded in the Atlas and a wide set of possible covariates, all with a number of levels (see the Appendix).

@@

Following @bourguignon1968, we reduce the complexity of the outcome variable to two levels: whether a society has possession trance or not. The predictors are the possible covariates identified above. Some predictors, like `EA032` (Jurisdictional hierarchy of local community) or `EA113` can be construed to follow a scale. These are input as ordered factors. Others, like `EA053` (Sex differences in animal husbandry) or `EA043` (Major type of descent) can be carved up into categories: These are one-hot encoded. For details, see the Appendix.

@@

```{r }
vars = ld %>% 
  filter(var_title != 'Trance states') %>% 
  distinct(var_id,var_title,var_definition,var_name) %>%
  arrange(var_id) %>% 
  group_by(var_id, var_title, var_definition) %>% 
  nest() %>% 
  mutate(levels = map(data, ~ paste(.$var_name, collapse = ', '))) %>% 
  select(var_id, var_title, levels) %>% 
  unnest(cols = c(levels)) %>% 
  rename(ID = var_id, Name = var_title)

varimp1 = read_tsv('models/varimp11.tsv')
vars2 = varimp1 %>% 
  slice(1:10) %>% 
  mutate(ID = str_extract(variable, '^EA...')) %>% 
  left_join(vars, by = "ID") %>% 
  select(ID, Name, relative_importance)

@@

## Appendix

### Relevant predictors in the Ethnographic Atlas

A survey of the literature resulted in the following predictors of trance and possession phenomena.

```{r }
kable(vars, 'simple')
```

### Recoded values in the Ethnographic Atlas

The recoded values were used in the model for each predictor. Where the recoded values are numeric, the predictor was used as an ordered factor. Where the recoded values are strings, the predictor was one-hot encoded (such that predictor A with levels 1, 2, and 3 was turned into boolean predictors A==2 and A==3).

```{r }
ld %>% 
  filter(var_id != 'EA112') %>% 
  distinct(var_id,var_title,var_description,recoded_value) %>% 
  arrange(var_id,recoded_value) %>% 
  kable('simple')
```
